
\section{Acknowledgements}

This work was supported by National Science Foundation (NSF) awards: ``Geoinformatics: A Petascale Cyberfacility for Physics-Based Seismic Hazard Analysis (SCEC PetaSHA3 Project)'' (EAR-0949443); ``SI2-SSI: A Sustainable Community Software Framework for Petascale Earthquake Modeling'' (ACI-1148493); ``Community Computational Platforms for Developing Three-Dimensional Models of Earth Structure'' (EAR-1226343); and ``Community Computational Platforms for Developing Three-Dimensional Models of Earth Structure, Phase II'' (EAR-1349180). SCEC is funded by NSF Cooperative Agreement EAR-1033462 and U.S.~Geological Survey Cooperative Agreement G12AC20038. The SCEC contribution number for this paper is 2067. This research is also part of the Blue Waters sustained-petascale computing project, which is supported by NSF (awards OCI-0725070 and ACI-1238993) and the state of Illinois. Blue Waters is a joint effort of the University of Illinois at Urbana-Champaign and its National Center for Supercomputing Applications (NCSA). Computational support was posible through PRAC allocations supported by NSF awards: ``Petascale Research in Earthquake System Science on Blue Waters (PressOnBlueWaters)'' (NSF ACI-0832698); and ``Extending the Spatiotemporal Scales of Physics-Based Seismic Hazard Analysis'' (ACI-1440085). Additional computational support was provided by the Extreme Science and Engineering Discovery Environment (XSEDE) program, which is supported by NSF (ACI-1053575). Computations referred here were performed on Kraken at the U.S.~National Institute for Computational Sciences (NICS) and Blue Waters at NCSA. We want to extend a special acknowledgement to community velocity modelers Harold Magistrale, John Shaw, Andreas Plesch, Peter S\"{u}ss, Carl Tape, Guoqing Lin, En-Jui Lee, Po Chen, Robert W.~Graves, and many other model contributors. UCVM would not be possible without your models and your input at various development stages during the years has been invaluable. We also want to thank the Carnegie Mellon etree library development group led by David O'Hallaron for their multiple direct and indirect contributions to the development of UCVM.
