\input{introduction}\section{Community Velocity Models}\label{sec:cvms}\textit{\color{blue}This section will present how CVMs work and the various CVMs available to the community today. It will basically explain that CVMs provide the triplets of Vs, Vp and density, and, as an example, we can expand on a description of CVM-S and CVM-H, including their variations CVM-SI and CVM-H+GTL.}\section{The UCVM Software Framework}\label{sec:ucvm}\textit{\color{blue}This section will describe in detail the various software components of UCVM. The list of sub-sections follows the main items in the UCVM user guide currently available in the Wiki. They are a tentative idea, but we should probably have a discussion on how to best lay out the different parts.}\subsection{Concept of Operation}\subsection{Utilities}\subsubsection{Querying}\subsubsection{Gridding and Meshing}\subsubsection{UCVM and the Etree Library}\subsubsection{Parallel Utilities}\subsection{Application Programming Interface}\section{Computational Performance}\label{sec:conclusions}\textit{\color{blue}It would be desirable to do some experiments in terms of performance, especially for the parallel applications utilities in UCVM. With some examples of how long the same thing would take if doing it differently. We will need to discuss this.}\section{Recent Case Applications}\label{sec:conclusions}\textit{\color{blue}This section will be dedicated to show case applications. Some ideas for potential subsections follow.}\subsection{Visualization and Model Comparisons}\subsection{Chino Hills Simulation Series}\subsection{CyberShake}\section{Summary and Conclusions}\label{sec:conclusions}\textit{\color{blue}A couple of paragraphs with a summary of what is shown in the paper and a few key conclusions about the impact that we expect UCVM has already have and will have on earthquake research.}