
\input{table-cvms}

\input{figure-cvms}

\subsection{Supported Velocity Models and Datasets}
\label{sec:cvms}

Existing geological velocity models vary considerably in terms of area of coverage, depth extent, composition, and resolution. The UCVM framework is flexible in its support for such variability and has been designed to integrate different velocity models, as well as to interact with their individual features seamlessly. UCVM has built-in support for a number of standard CVMs, which UCVM utilities identify through a series of corresponding string labels. Table \ref{tab:cvms} lists the models currently supported in the UCVM framework and provides additional details and references, along with their string labels. While all these models are supported by UCVM, only a fraction of them are included in the automated installation package. The remaining ones need to be installed manually before they can be accessed through the platform. Additional details are given in the documentation provided in Table \ref{tab:manuals} and in Section \ref{sec:easy.install}. In Table \ref{tab:cvms}, we indicate which of the models are included in the automated installation and which require manual installation. Figure \ref{fig:cvms} shows a map with the coverage areas of various CVMs.

UCVM is distributed with an easy installation method which runs a Python script (\texttt{ucvm\_set\-up.py}) that prompts the user about which of the automated models are to be installed (see Section \ref{sec:easy.install}). Before they are accessible through the UCVM framework, velocity models need to be enabled through a UCVM configuration file entry, and UCVM needs to be compiled. If the user wants to enable other velocity models at a later time, the framework needs to be compiled again in order to enable the desired additional models. Advanced users can also customize the installation to support other models, including user-defined velocity models, which can be added in the form of an \textit{etree} \citep{Tu_2003_Tech} database or as a \textit{patch} (rasterized) model. These advanced features and other options mentioned here are described in greater detail in the UCVM Advanced User guide (see Table \ref{tab:manuals}), and are controlled at run-time through a configuration file read by UCVM to identify the model(s) to be queried and the different alternative features.

In addition to standard velocity models, UCVM includes two geotechnical layer (GTL) models, also listed in Table \ref{tab:cvms}. These models are intended to supplement (i.e., modify or replace) the near-surface information in the original velocity models for a smoother transition from the softer near-surface soil deposits to the stiffer bedrock basement. The first of the two GTL models implements a \vsthirty-based interpolation from the free surface down to a given depth $z$, following \citet{Ely_2010_AGU}. The second GTL model is a generic one-dimensional (1D) model identical to the 1D crustal model. Two interpolation schemes can be used to smooth GTL material properties with the underlying crustal model material properties: a linear interpolation, or the interpolation relationship used by \citet{Ely_2010_AGU}. As in the case of the CVMs, the GTLs and the interpolation schemes are identified with string labels, \texttt{elygtl} and \texttt{1dgtl} for the \vsthirty-based and the 1D models, respectively. Similarly, the interpolation schemes are identified with the labels \texttt{ely} and \texttt{linear}. Note that the Ely GTL and interpolation models are used in the CVM-H model by default, with a reference depth of $z=350$~m. This option can be turned on or off for CVM-H by setting the model flag \texttt{USE\_GTL} \texttt{=} \texttt{true/false} in the configuration file.

To support the \vsthirty-based GTL model, UCVM has two built-in standard maps for California at 1 arcsec resolution (following USGS NED standards). These maps contain elevation data and \vsthirty{} data for the region indicated in Figure \ref{fig:cvms}. The first standard map contains a composite \vsthirty{} dataset derived from the work of \citet{Wills_2006_BSSA} and \citet{Wald_2007_BSSA}, while the second provides an alternative dataset derived from \citet{Yong_2012_BSSA}. The choice of map is selectable by the user, with the first option identified by the string label \texttt{ucvm}, which is the default option, and \texttt{yong} for the latter. These maps are stored as etree databases, and their details are included in Table \ref{tab:cvms} and Figure \ref{fig:cvms}.

Since most models are bounded to a particular coverage area, UCVM provides alternative options to support background models. The CVM-S model, for instance, reverts to a 1D velocity model outside its defined region (see references provided in Table \ref{tab:cvms}). The CVM-H model, on the other hand, is supported through UCVM with an optional background 1D velocity structure. This option allows queries at points outside and below the domain covered by the standard version of the model. While this option is inactive by default, it can be controlled manually by setting the model flag \texttt{USE\_1D\_BKG} \texttt{=} \texttt{true}. Other models can make use of this background structure by including the 1D model in the list of query models; that is, in the models sequence used to compose the meta-model. We refer to this process as tiling and explain its implementation in Section \ref{sec:querying}.

